---
title: "Current Letter"
format: 
  html:
    embed-resources: true
---

```{r}
here::i_am("CurrentLetter.qmd")
library(here)
library(tidyverse)
library(ggrepel)
library(gghighlight)
library(ggpubr)
library(scales)
```


```{r, include=FALSE}
source(here("read_and_process_data.R"))
```



# Teacher salaries vs per capita income 
Considering Massachusetts municipalities with their own k-12 school districts 
(as opposed to those participating in regional school districts)
only `r label_percent(accuracy=1)(1-mhd_3_way_percentiles$pci_percentile)` 
of these municipalities have per capita income higher than Marblehead, while a full 
`r label_percent(accuracy=1)(1-mhd_3_way_percentiles$salary_percentile)` 
of them have a higher average teacher salary.


# Per Captia Income and Teacher Average Salary


Less than `r label_percent(accuracy=1)(1-mhd_percentiles$pci_percentile)` 
of Massachusetts municipalities have per capita income higher than Marblehead, while a full 
`r label_percent(accuracy=1)(1-mhd_inc_sal_percentiles$salary_percentile)` 
of school districts in the state have a higher average teacher salary.

```{r}
mhd_pc_receipts <- rev_inc %>%
  filter(Municipality=="Marblehead") %>%
  pull(per_capita_receipts)

peer_median_pc_receipts <- rev_inc %>%
  filter(pci_peer) %>%
  summarize(median_pc_receipts = median(per_capita_receipts)) %>%
  pull(median_pc_receipts)
```

# Tax receipts vs Per Capita Income
```{r}
n_municipalities <- length(levels(rev_inc$Municipality))
```

There's been plenty of discussion in Marblehead lately about the town's finances.
Overrides have been voted down, as have zoning density increases that could
bring growth to the property tax base.

With the apparent reluctance on the part of voters to raise
new revenue, and multiple town departments staring down the need for cuts in services to balance the budget, we wondered how Marblehead's tax burden compared to its peers.

The state data on taxes and income covers `r n_municipalities` municipalities.

We looked at per capita tax receipts (Department of Revenue data for Total Receipts excluding Enterprise and CPA funds),
and compared it to per capita income.


Marblehead ranks `r max(rev_inc$pci_rank) - mhd_percentiles$pci_rank + 1` in per captia income among the `r n_municipalities` municipalities in the DOR data. In terms of tax receipts per capita, Marblehead ranks
`r max(rev_inc$pc_receipts_rank) - mhd_percentiles$pc_receipts_rank + 1` of Massachusetts municipalities.

To create a peer group for comparison, we selected the 10% of Mass municipalities closest to Marblehead in the per capita income rankings, giving us a list of `r nrow(rev_inc %>% filter(pci_peer))` municipalities, with Marbhelehad right in the middle.

Marblehead's per capita tax receipts of  `r label_dollar()(mhd_pc_receipts)` is
`r label_percent()((peer_median_pc_receipts - mhd_pc_receipts)/peer_median_pc_receipts)` 
lower than the median of `r label_dollar()(peer_median_pc_receipts)` of its 
closest per capita income peers.


