---
title: "MHD Finances"
format: 
  html:
    embed-resources: true
---

```{r}
here::i_am("MHDRevenue.qmd")
library(here)
library(tidyverse)
library(readxl)
library(scales)
```
```{r, include=FALSE}
source(here("read_and_process_data.R"))
```


```{r}
n_municipalities <- length(levels(rev_inc$Municipality))
```

The state data on taxes and incomes covers `r n_municipalities` municipalities.

```{r}
mhd_percentiles
```


Marblehead ranks
`r max(rev_inc$pci_rank) - mhd_percentiles$pci_rank + 1` 
in income per capita of the `r n_municipalities` municipalities in the state, higher than `r label_percent(accuracy=0.1)(mhd_percentiles$pci_percentile)` of them.

In terms of tax receipts per capita, Marblehead ranks
`r max(rev_inc$pc_receipts_rank) - mhd_percentiles$pc_receipts_rank + 1` of Massachusetts municipalities.


```{r}
rev_inc %>%
  ggplot(aes(x=DOR.Income.Per.Capita, y=per_capita_receipts, label=Municipality)) +
  geom_point() +
  gghighlight::gghighlight(Municipality == "Marblehead") +
  scale_x_continuous(labels=label_dollar()) +
  scale_y_continuous(labels=label_dollar()) +
  labs(title="All Massachusetts Municipalities",
         x="Income per Capita",
         y="Per Capita Tax Receipts")
```




```{r}
mhd_pc_receipts <- rev_inc %>%
  filter(Municipality=="Marblehead") %>%
  pull(per_capita_receipts)

peer_median_pc_receipts <- rev_inc %>%
  filter(pci_peer) %>%
  summarize(median_pc_receipts = median(per_capita_receipts)) %>%
  pull(median_pc_receipts)
```



# Marblehead and its per capita income peers


By comparing to a peer group of the 10% of Massachusetts municipalities with
income per capita closest to Marblehead's, we find that
Marblehead's per capita tax receipts of  `r label_dollar()(mhd_pc_receipts)` is
`r label_percent()((peer_median_pc_receipts - mhd_pc_receipts)/peer_median_pc_receipts)` 
lower than the median of `r label_dollar()(peer_median_pc_receipts)` of its 
closest per capita income peers.


```{r}
rev_inc %>%
  filter(pci_peer) %>%
  mutate(class = case_when(
    Municipality=="Marblehead" ~ "Marblehead",
    ns_peer ~ "Essex County",
    TRUE ~ "Other Mass Municipalities"
  ),
  label=case_when(!ns_peer ~ NA,
                  TRUE ~ Municipality)) %>%
  ggplot(aes(x=DOR.Income.Per.Capita, y=per_capita_receipts, label=Municipality,
             color=class)) +
  geom_point() +
  ggrepel::geom_text_repel(show.legend = FALSE) +
  geom_hline(yintercept=peer_median_pc_receipts, linetype="dashed",
             color="#333333") +
  annotate(geom="text", x=140e3, y=peer_median_pc_receipts+150, label="Median of Peers",
           color="#333333"
           ) +
  scale_color_manual(values=c("Marblehead"="#F8766D", "Essex County"="#00BFC4", "Other Mass Municipalities"="gray"))+
  scale_x_continuous(labels=label_dollar()) +
  scale_y_continuous(labels=label_dollar()) +
  labs(title="Tax Receipts among Marblehead's Per Capita Income Peers",
         x="Income per Capita",
         y="Per Capita Tax Receipts") +
  theme_minimal() +
  theme(legend.title = element_blank(),
        legend.position = "bottom")
```


```{r}
rev_inc %>%
  filter(pci_peer) %>%
  select(Municipality, 
         ns_peer,
         DOR.Income.Per.Capita,
         per_capita_receipts,
         pc_levy,
         contains("rank"), contains("percentile")) %>%
  rename(`Essex County` = ns_peer) %>%
  DT::datatable(filter="top", rownames=FALSE)
```



```{r}
rev_inc %>%
  filter(pci_peer) %>%
  #calculate mean and median of DOR.Income.Per.Capita and per_capita_receipts
  summarize(mean_income = mean(DOR.Income.Per.Capita),
            median_income = median(DOR.Income.Per.Capita),
            mean_receipts = mean(per_capita_receipts),
            median_receipts = median(per_capita_receipts))
```

# among k-12 district peers
```{r}
rev_inc_sal %>%
  ggplot(aes(x=DOR.Income.Per.Capita, y=per_capita_receipts, label=Municipality)) +
  geom_point() +
  geom_smooth(method="loess", se=FALSE) +
  gghighlight::gghighlight(Municipality == "Marblehead") +
  scale_x_continuous(labels=label_dollar()) +
  scale_y_continuous(labels=label_dollar()) +
  labs(title="Mass municipalities with their own k-12 school districts",
         x="Income per Capita",
         y="Per Capita Tax Receipts")
```

```{r}
k_12_peer_median_pc_receipts <- rev_inc_sal %>%
  filter(pci_peer) %>%
  summarize(median_pc_receipts = median(per_capita_receipts)) %>%
  pull(median_pc_receipts)
```



# Marblehead and its per capita income peers


By comparing to a peer group of the 10% of Massachusetts municipalities with
income per capita closest to Marblehead's, we find that
Marblehead's per capita tax receipts of  `r label_dollar()(mhd_pc_receipts)` is
`r label_percent()((k_12_peer_median_pc_receipts - mhd_pc_receipts)/k_12_peer_median_pc_receipts)` 
lower than the median of `r label_dollar()(k_12_peer_median_pc_receipts)` of its 
closest per capita income peers.


```{r, fig.height=5, fig.width=5}
plot_rev_income <- rev_inc_sal %>%
  filter(pci_peer) %>%
  mutate(class = case_when(
    Municipality=="Marblehead" ~ "Marblehead",
    ns_peer ~ "Essex County",
    TRUE ~ "Other Mass Municipalities"
  ),
  label=case_when(!ns_peer ~ NA,
                  TRUE ~ Municipality)) %>%
  ggplot(aes(x=DOR.Income.Per.Capita, y=per_capita_receipts, label=Municipality,
             color=class)) +
  geom_point() +
  ggrepel::geom_text_repel(show.legend = FALSE) +
  geom_hline(yintercept=k_12_peer_median_pc_receipts, linetype="dashed",
             color="#333333") +
  annotate(geom="text", x=120e3, y=k_12_peer_median_pc_receipts-100, label="Median of Peers",
           color="#333333"
           ) +
  scale_color_manual(values=c("Marblehead"="#F8766D", "Essex County"="#00BFC4", "Other Mass Municipalities"="gray"))+
  scale_x_continuous(labels=label_dollar()) +
  scale_y_continuous(labels=label_dollar()) +
  labs(title="Tax Receipts",
       subtitle="among Marblehead's Per Capita Income Peers",
         x="Income per Capita",
         y="Per Capita Tax Receipts") +
  theme_minimal() +
  theme(legend.title = element_blank(),
        legend.position = "bottom")
ggsave(here("figures", "rev_vs_income.png"), plot_rev_income, width = 5, height = 5, units = "in")
plot_rev_income
```
# barplots of income and tax receipts
```{r}
pci_bar_plot <- rev_inc_sal %>%
  filter(pci_peer) %>%
  mutate(Municipality=fct_reorder(Municipality, DOR.Income.Per.Capita)) %>%
  mutate(FillCol = ifelse(Municipality == "Marblehead", "#00BFC4", "#F8766D")) %>%
  ggplot(aes(y = Municipality, x = DOR.Income.Per.Capita, fill = FillCol)) +
  geom_col() +
  theme_minimal() +
  labs(title = "Per Capita Income",
       x = "DOR Income Per Capita") +
  geom_text(aes(label = scales::dollar(DOR.Income.Per.Capita)), hjust = -0.1, size = 3)+
  guides(fill="none")+
  coord_cartesian(xlim = c(0, 260000)) +
  theme(plot.caption = element_text(color = "white"),axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(), 
        axis.title=element_blank(),
        axis.text=element_text(size=10),
        title=element_text(size=10))

pci_bar_plot
```
# Average teacher salary of Marblehead and its income peers

```{r}
receipts_bar_plot <- rev_inc_sal %>%
  filter(pci_peer) %>%
  mutate(Municipality=fct_reorder(Municipality, per_capita_receipts),
         label=label_dollar(accuracy=1)(per_capita_receipts)) %>%
  mutate(FillCol = ifelse(Municipality == "Marblehead", "#00BFC4", "#F8766D")) %>%
  ggplot(aes(y = Municipality, x = per_capita_receipts, fill = FillCol, label=label)) +
  geom_col() +
  theme_minimal() +
  labs(title = "Per Capita Tax Receipts",
       y = "",
       x = "Per Captia Tax Receipts",
       caption = "*Districts with per capita income \nsimilar to Marblehead") +
  geom_text(hjust = -0.1, size = 3)+
   guides(fill="none")+
   coord_cartesian(xlim = c(0, 10e3)) +
  theme(axis.title=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.text=element_text(size=10),
        title=element_text(size=10))

receipts_bar_plot
```

# Composite plot, per captia income and average teacher salary
```{r}
comp <- plot_grid(pci_bar_plot, receipts_bar_plot, align = "h")
ggsave(here("figures", "pci_receipts_comp.png"), comp, width = 6, height = 4, units = "in")
comp
```

```{r}
rev_inc %>%
  filter(County=="ESSEX") %>%
  mutate(class = case_when(
    Municipality=="Marblehead" ~ "Marblehead",
    TRUE ~ "Others"
  ),
  label=case_when(class=="Others" ~ NA,
                  TRUE ~ Municipality)) %>%
  ggplot(aes(x=DOR.Income.Per.Capita, y=per_capita_receipts, label=label,
             color=class)) +
  geom_point() +
  #ggrepel::geom_label_repel() +
  scale_x_continuous(labels=label_dollar()) +
  scale_y_continuous(labels=label_dollar()) +
  labs(title="Essex County Municipalities",
         x="Income per Capita",
         y="Per Capita Tax Receipts")
```


# Sources
Revenue by source: https://dlsgateway.dor.state.ma.us/reports/rdPage.aspx?rdReport=RevenueBySource.RBS.RevbySourceMAIN

Income per capita : https://dlsgateway.dor.state.ma.us/reports/rdPage.aspx?rdReport=DOR_Income_EQV_Per_Capita

