---
title: "MHD Finances"
format: 
  html:
    embed-resources: true
---

```{r}
here::i_am("MHDRevenue.qmd")
library(here)
library(tidyverse)
library(readxl)
library(scales)
```
```{r, include=FALSE}
source(here("read_and_process_data.R"))
```


```{r}
n_municipalities <- length(levels(rev_inc$Municipality))
```

The state data on taxes and incomes covers `r n_municipalities` municipalities.

```{r}
mhd_percentiles
```


Marblehead ranks
`r max(rev_inc$pci_rank) - mhd_percentiles$pci_rank + 1` 
in income per capita of the `r n_municipalities` municipalities in the state, higher than `r label_percent(accuracy=0.1)(mhd_percentiles$pci_percentile)` of them.

In terms of tax receipts per capita, Marblehead ranks
`r max(rev_inc$pc_receipts_rank) - mhd_percentiles$pc_receipts_rank + 1` of Massachusetts municipalities.

Less than `r label_percent(accuracy=1)(1-mhd_percentiles$pci_percentile)` 
of Massachusetts municipalities have per capita income higher than Marblehead, while a full 
`r label_percent(accuracy=1)(1-mhd_inc_sal_percentiles$salary_percentile)` 
of school districts in the state have a higher average teacher salary.


```{r}
rev_inc %>%
  ggplot(aes(x=DOR.Income.Per.Capita, y=per_capita_receipts, label=Municipality)) +
  geom_point() +
  gghighlight::gghighlight(Municipality == "Marblehead") +
  scale_x_continuous(labels=label_dollar()) +
  scale_y_continuous(labels=label_dollar()) +
  labs(title="All Massachusetts Municipalities",
         x="Income per Capita",
         y="Per Capita Tax Receipts")
```

```{r}
mhd_pc_receipts <- rev_inc %>%
  filter(Municipality=="Marblehead") %>%
  pull(per_capita_receipts)

peer_median_pc_receipts <- rev_inc %>%
  filter(pci_peer) %>%
  summarize(median_pc_receipts = median(per_capita_receipts)) %>%
  pull(median_pc_receipts)
```



# Marblehead and its per capita income peers


By comparing to a peer group of the 10% of Massachusetts municipalities with
income per capita closest to Marblehead's, we find that
Marblehead's per capita tax receipts of  `r label_dollar()(mhd_pc_receipts)` is
`r label_percent()((peer_median_pc_receipts - mhd_pc_receipts)/peer_median_pc_receipts)` 
lower than the median of `r label_dollar()(peer_median_pc_receipts)` of its 
closest per capita income peers.


```{r}
rev_inc %>%
  filter(pci_peer) %>%
  mutate(class = case_when(
    Municipality=="Marblehead" ~ "Marblehead",
    ns_peer ~ "North Shore",
    TRUE ~ "Other Mass Municipalities"
  ),
  label=case_when(!ns_peer ~ NA,
                  TRUE ~ Municipality)) %>%
  ggplot(aes(x=DOR.Income.Per.Capita, y=per_capita_receipts, label=label,
             color=class)) +
  geom_point() +
  ggrepel::geom_text_repel() +
  geom_hline(yintercept=peer_median_pc_receipts, linetype="dashed",
             color="darkgray") +
  annotate(geom="text", x=140e3, y=peer_median_pc_receipts+150, label="Median of Peers",
           color="darkgray"
           ) +
  scale_x_continuous(labels=label_dollar()) +
  scale_y_continuous(labels=label_dollar()) +
  labs(title="Marblehead's Per Capita Income Peers",
       subtitle = "the 34 Mass municipalities (17 above, 17 below) closest to Marblehead in income per capita",
         x="Income per Capita",
         y="Per Capita Tax Receipts") +
  theme_minimal()
```


```{r}
rev_inc %>%
  filter(pci_peer) %>%
  select(Municipality, 
         ns_peer,
         DOR.Income.Per.Capita,
         per_capita_receipts,
         pc_levy,
         contains("rank"), contains("percentile")) %>%
  rename(`Essex County` = ns_peer) %>%
  DT::datatable(filter="top", rownames=FALSE)
```



```{r}
rev_inc %>%
  filter(pci_peer) %>%
  #calculate mean and median of DOR.Income.Per.Capita and per_capita_receipts
  summarize(mean_income = mean(DOR.Income.Per.Capita),
            median_income = median(DOR.Income.Per.Capita),
            mean_receipts = mean(per_capita_receipts),
            median_receipts = median(per_capita_receipts))
```


```{r}
rev_inc %>%
  filter(County=="ESSEX") %>%
  mutate(class = case_when(
    Municipality=="Marblehead" ~ "Marblehead",
    TRUE ~ "Others"
  ),
  label=case_when(class=="Others" ~ NA,
                  TRUE ~ Municipality)) %>%
  ggplot(aes(x=DOR.Income.Per.Capita, y=per_capita_receipts, label=label,
             color=class)) +
  geom_point() +
  #ggrepel::geom_label_repel() +
  scale_x_continuous(labels=label_dollar()) +
  scale_y_continuous(labels=label_dollar()) +
  labs(title="Essex County Municipalities",
         x="Income per Capita",
         y="Per Capita Tax Receipts")
```


# Sources
Revenue by source: https://dlsgateway.dor.state.ma.us/reports/rdPage.aspx?rdReport=RevenueBySource.RBS.RevbySourceMAIN

Income per capita : https://dlsgateway.dor.state.ma.us/reports/rdPage.aspx?rdReport=DOR_Income_EQV_Per_Capita

