---
title: "MHD schools data import and analysis"
format: 
  html:
    embed-resources: true
---


```{r}
here::i_am("read_and_process_data.qmd")
library(here)
library(tidyverse)
library(DT)
```

```{r}
source(here("read_and_process_data.R"))
```

# Districts with full k-12 enrollment
```{r}
enrollment %>%
  filter(full_k_12) %>%
  select(District_NAME, District_Total) %>%
  datatable(filter="top", rownames=FALSE)
```

# Municipalities with a school district matching their name
```{r}
rev_inc %>%
  filter(Municipality %in% k_12_districts) %>%
  datatable(filter="top", rownames=FALSE)
```



# municipalities in the teacher salary data that are not in the DOR data
```{r}
#find the sal that are not in inc
sal[!(sal$Municipality %in% inc$Municipality),] %>%
  arrange(desc(FTE.Count)) %>%
  datatable(filter="top", rownames = FALSE)
```


# Municipalities dropped from the DOR data for missing teacher salary data
```{r}
rev_inc %>%
  filter(Municipality %in% setdiff(rev_inc$Municipality, rev_inc_sal$Municipality)) %>%
  datatable(filter="top", rownames=FALSE)
```

NOTE we can get some of these back by assigning the municipality to their 
appropriate regional school district e.g. Hamilton and Wehnam to Hamilton-Wenham


# Sources
district enrollment by grade https://www.doe.mass.edu/infoservices/reports/enroll/default.html?yr=2324
