---
title: "School Salary"
format: 
  html:
    embed-resources: true
---

```{r}
here::i_am("SchoolSalary.qmd")
library(here)
library(readxl)
library(tidyverse)
library(ggrepel)
library(gghighlight)
library(ggpubr)
library(scales)
library(cowplot)
library(DT)
```

```{r, include=FALSE}
source(here("read_and_process_data.R"))
```


# Educators by age group
```{r}
#plot age_long AgeGroup vs Count using ggplot
plot1 <- ggplot(age_long, aes(x = AgeGroup, y = Percent)) +
  geom_point(alpha = .01) +
  theme_minimal() +
  guides(fill="none")+
  #geom_point(data = age_long %>% filter(District.Name == "Marblehead"), aes(x = AgeGroup, y = Percent, color = ""#F8766D""))+
  geom_label(data = age_long %>% filter(District.Name == "Marblehead"), aes(label = round(Percent, digits = 2), x = AgeGroup, y = Percent), color = "#F8766D")+
  labs(x = "Age Group",
       y = "Proportion of Educators")+
    theme(legend.position="none")

ggsave(here("figures", "age.png"), plot1, width = 4, height = 2, units = "in")

plot1
```


# Average teacher salary vs per capita income

Using only k-12 districts

```{r}
plot2 <- ggplot(rev_inc_sal, aes(x = DOR.Income.Per.Capita, y = Average.Salary)) +
  geom_point(alpha = .5) +
  geom_smooth(method = "loess", se = FALSE, color = "#7CAE00") +
  theme_minimal()+
  geom_text_repel(data = rev_inc_sal %>% filter( !(Municipality == "Marblehead")), aes(label = Municipality), max.overlaps = 20, size = 3, segment.color = 'grey80') +
  geom_label_repel(data = rev_inc_sal %>% filter(Municipality == "Marblehead"), aes(label = Municipality), color = "#F8766D", position = position_nudge_repel(x =100, y =15))+
  geom_label_repel(data = rev_inc_sal %>% filter(Municipality == "Gloucester" |Municipality == "Beverly"), aes(label = Municipality), color = "#3B9AB2") +
  geom_point(data = rev_inc_sal %>% filter(Municipality == "Marblehead"), color = "#F8766D", size = 3) +
  geom_point(data = rev_inc_sal %>% filter(Municipality == "Gloucester" |Municipality == "Beverly"), color = "#00BFC4", size = 2) +
  scale_x_continuous(labels = scales::label_dollar()) +
  scale_y_continuous(labels = scales::label_dollar())+
  labs(x = "Per Capita Income",
       y = "Average Teacher Salary", 
       caption = "*DOR Income Per Capita represents the annual income as \nreported on annual tax returns divided by the US Census Population.")+
  theme(plot.caption.position = "plot",
        plot.caption = element_text(hjust = 0),
        axis.text=element_text(size=12),
        axis.title=element_text(size=18))
ggsave(here("figures", "sal.png"), plot2, width = 9, height = 9, units = "in")
plot2
```

```{r, warning=FALSE}
plot_sal_pci <- rev_inc_sal %>%
  mutate(plot_class = case_when(Municipality == "Marblehead" ~ "MHD",
                                ns_peer ~ "ns_peer",
                                TRUE ~ "other"),
         label=case_when(plot_class != "other" ~ Municipality,
                         TRUE ~ NA)) %>%
  ggplot(aes(x=DOR.Income.Per.Capita, y=Average.Salary, label=label, color = plot_class)) +
  geom_point() + 
  geom_smooth(method = "loess", se = FALSE, color = "#7CAE00") +
  geom_text_repel(data=rev_inc_sal %>% filter(!ns_peer),
                  aes(x=DOR.Income.Per.Capita, y=Average.Salary, label=Municipality), color = "grey") +
  geom_label_repel(data=rev_inc_sal %>% filter(ns_peer & !Municipality=="Marblehead"),
                  aes(x=DOR.Income.Per.Capita, y=Average.Salary, label=Municipality), color = "#00BFC4") +
  annotate(geom="label",
           x=rev_inc_sal %>% filter(Municipality == "Marblehead") %>% pull(DOR.Income.Per.Capita) + 34e3,
           y=rev_inc_sal %>% filter(Municipality == "Marblehead") %>% pull(Average.Salary) - 2e3,
           label="Marblehead", color = "#F8766D") +
  # geom_label_repel() +
  scale_color_manual(values = c(MHD="#F8766D", ns_peer="#00BFC4", other="grey")) +
  scale_x_continuous(labels = scales::label_dollar()) +
  scale_y_continuous(labels = scales::label_dollar()) +
  theme_minimal()  +
  theme(legend.position="none") + 
  labs(x = "Per Capita Income",
       y = "Average Teacher Salary", 
       caption = "*DOR Income Per Capita represents the annual income as \nreported on tax returns divided by the US Census Population.")

ggsave(here("figures", "sal_vs_pci.png"), plot_sal_pci, width = 6, height = 4, units = "in")

plot_sal_pci
```
```{r}
rev_inc_sal %>%
  mutate(plot_class = case_when(Municipality == "Marblehead" ~ "MHD",
                                ns_peer ~ "ns_peer",
                                TRUE ~ "other"),
         label=case_when(plot_class != "other" ~ Municipality,
                         TRUE ~ NA)) %>%
  ggplot(aes(x=DOR.Income.Per.Capita, y=Average.Salary, label=label, color = plot_class)) +
  geom_point() + 
  geom_smooth(method = "loess", se = FALSE, color = "#7CAE00") +
  geom_text_repel(data=rev_inc_sal %>% filter(!ns_peer),
                  aes(x=DOR.Income.Per.Capita, y=Average.Salary, label=Municipality), color = "grey") +
  geom_label_repel(data=rev_inc_sal %>% filter(ns_peer & !Municipality=="Marblehead"),
                  aes(x=DOR.Income.Per.Capita, y=Average.Salary, label=Municipality), color = "#00BFC4") +
  annotate(geom="label",
           x=rev_inc_sal %>% filter(Municipality == "Marblehead") %>% pull(DOR.Income.Per.Capita) + 5e3,
           y=rev_inc_sal %>% filter(Municipality == "Marblehead") %>% pull(Average.Salary) - 2e3,
           label="Marblehead", color = "#F8766D") +
  # geom_label_repel() +
  scale_color_manual(values = c(MHD="#F8766D", ns_peer="#00BFC4", other="grey")) +
  scale_x_continuous(labels = scales::label_dollar()) +
  scale_y_continuous(labels = scales::label_dollar()) +
  theme_minimal()  +
  theme(legend.position="none") + 
  labs(x = "Per Capita Income",
       y = "Average Teacher Salary", 
       caption = "*DOR Income Per Capita represents the annual income as \nreported on tax returns divided by the US Census Population.") +
  coord_cartesian(xlim = c(70e3, 120e3), ylim = c(70e3, 110e3))
```


```{r}
rev_inc_sal %>%
  filter(pci_peer) %>%
  mutate(plot_class = case_when(Municipality == "Marblehead" ~ "MHD",
                                ns_peer ~ "ns_peer",
                                TRUE ~ "other"),
         label=case_when(plot_class != "other" ~ Municipality,
                         TRUE ~ Municipality)) %>%
  ggplot(aes(x=DOR.Income.Per.Capita, y=Average.Salary, label=label, color = plot_class)) +
  geom_point() + 
  geom_text_repel() +
  scale_color_manual(values = c(MHD="#F8766D", ns_peer="#00BFC4", other="grey")) +
  scale_x_continuous(labels = scales::label_dollar()) +
  scale_y_continuous(labels = scales::label_dollar()) +
  theme_minimal() +
  theme(legend.position="none") + 
  labs(x = "Per Capita Income",
       y = "Average Teacher Salary", 
       caption = "*DOR Income Per Capita represents the annual income as \nreported on tax returns divided by the US Census Population.")
```

```{r}
rev_inc_sal %>%
  mutate(plot_class = case_when(
    Municipality == "Marblehead" ~ "Marblehead",
    ns_peer ~ "North Shore PCI Peers",
    TRUE ~ "Others"),
    label=case_when(Municipality == "Marblehead" ~ "Marblehead",
                    ns_peer ~ Municipality,
                    TRUE ~ NA)) %>%
  select(Municipality, contains("peer"), plot_class, label) %>%
  datatable(filter="top", rownames=FALSE)
```


```{r}
mhd_income_peers %>%
  arrange(desc(Average.Salary)) %>%
  select(Municipality, Average.Salary, contains("percentile"))
```


# Per capita income of Marblehead and its income peers
```{r}
pci_bar_plot <- rev_inc_sal %>%
  filter(pci_peer) %>%
  mutate(Municipality=fct_reorder(Municipality, DOR.Income.Per.Capita)) %>%
  mutate(FillCol = ifelse(Municipality == "Marblehead", "#00BFC4", "#F8766D")) %>%
  ggplot(aes(y = Municipality, x = DOR.Income.Per.Capita, fill = FillCol)) +
  geom_col() +
  theme_minimal() +
  labs(title = "Per Capita Income",
       x = "DOR Income Per Capita") +
  geom_text(aes(label = scales::dollar(DOR.Income.Per.Capita)), hjust = -0.1, size = 3)+
  guides(fill="none")+
  coord_cartesian(xlim = c(0, 260000)) +
  theme(plot.caption = element_text(color = "white"),axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(), 
        axis.title=element_blank(),
        axis.text=element_text(size=10),
        title=element_text(size=10))

pci_bar_plot
```
# Average teacher salary of Marblehead and its income peers

```{r}
sal_bar_plot <- rev_inc_sal %>%
  filter(pci_peer) %>%
  mutate(Municipality=fct_reorder(Municipality, Average.Salary)) %>%
  mutate(FillCol = ifelse(Municipality == "Marblehead", "#00BFC4", "#F8766D")) %>%
  ggplot(aes(y = Municipality, x = Average.Salary, fill = FillCol)) +
  geom_col() +
  theme_minimal() +
  labs(title = "Average Teacher Salary",
       y = "",
       x = "Average Teacher Salary",
       caption = "*Districts with per capita income \nsimilar to Marblehead") +
  geom_text(aes(label = scales::dollar(Average.Salary)), hjust = -0.1, size = 3)+
   guides(fill="none")+
   coord_cartesian(xlim = c(0, 150000)) +
  theme(axis.title=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.text=element_text(size=10),
        title=element_text(size=10))

sal_bar_plot
```

# Composite plot, per captia income and average teacher salary
```{r}
comp <- plot_grid(pci_bar_plot, sal_bar_plot, align = "h")
ggsave(here("figures", "comp.png"), comp, width = 6, height = 4, units = "in")
comp
```

# All Mass municipalities, tax receipts vs per capita income
```{r}
plot5 <- rev_inc_sal %>%
   mutate(per_capita_receipts = Total.Receipts...9/Population) %>%
   ggplot(aes(x=DOR.Income.Per.Capita, y=per_capita_receipts, label=Municipality)) +
   geom_point(colour = "darkgrey", alpha = .5) +
   geom_label_repel(data = rev_inc %>% filter(Municipality == "Marblehead"), aes(label = Municipality), color = "#F8766D", position = position_nudge_repel(x =100, y =15))+
   geom_point(data = rev_inc %>% filter(Municipality == "Marblehead"), color = "#F8766D", size = 3) +
   scale_x_continuous(labels=label_dollar()) +
   scale_y_continuous(labels=label_dollar())+ 
   theme_minimal()+
   labs(x= "Per Capita Income",
        y = "Per Capita Tax Receipts")+
   theme(plot.caption.position = "plot",
         plot.caption = element_text(hjust = 0),
         axis.text=element_text(size=12),
         axis.title=element_text(size=18))
 
 ggsave(here("figures", "tax.png"), plot5, width = 6, height = 3, units = "in")
 plot5
```

# Average teacher salary vs per-capita tax receipts
```{r}
plot6<-  rev_inc_sal %>% 
  filter(per_capita_receipts < 12000) %>%
   ggplot(aes(x=per_capita_receipts, y=Average.Salary, label=Municipality)) +
   geom_point(colour = "darkgrey", alpha = .5) +
  #geom_text_repel(data = rev_inc_sal %>% filter( !(Municipality == "Marblehead")), aes(label = Municipality), max.overlaps = 5, size = 3, segment.color = 'grey80') +
  geom_smooth(method = "loess", se = FALSE, color = "#7CAE00") +
  geom_label_repel(data = rev_inc_sal %>% filter(Municipality == "Marblehead"), aes(label = Municipality), color = "#F8766D")+
   geom_point(data = rev_inc_sal %>% filter(Municipality == "Marblehead"), color = "#F8766D", size = 3) +
   scale_x_continuous(labels=label_dollar()) +
   scale_y_continuous(labels=label_dollar())+ 
   theme_minimal()+
   labs(x= "Per Capita Tax Receipts",
        y = "Average Teacher Salary")+
   theme(plot.caption.position = "plot",
         plot.caption = element_text(hjust = 0),
         axis.text=element_text(size=12),
         axis.title=element_text(size=18))
       
ggsave(here("figures", "taxsal.png"), plot6, width = 7, height = 3.5, units = "in")
plot6
```

```{r}
mhd_tax_peers
```
# Average teacher salary of marblehead and its tax peers
```{r}
#make a bar graph of the average salary of Marblehead and its tax peers
plot7<- mhd_tax_peers %>%
  mutate(Municipality=fct_reorder(Municipality, Average.Salary), 
         FillCol = ifelse(Municipality == "Marblehead", "#00BFC4", "#F8766D")) %>%
  ggplot(aes(y = Municipality, x = Average.Salary, fill = FillCol)) +
  geom_col() +
  theme_minimal() +
  labs(title = "Average Teacher Salary",
       y = "",
       x = "Average Teacher Salary",
       caption = "*Districts with per capita tax receipts \nsimilar to Marblehead") +
  geom_text(aes(label = scales::dollar(Average.Salary)), hjust = -0.1, size = 3)+
  guides(fill="none")+
  coord_cartesian(xlim = c(0, 150000)) +
  theme(axis.title=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.text=element_text(size=10),
        title=element_text(size=10))
plot7
```

```{r}
#make a bar graph of the average salary of Marblehead and its tax peers
plot8<- mhd_tax_peers %>%
  mutate(Municipality=fct_reorder(Municipality, per_capita_receipts), 
         FillCol = ifelse(Municipality == "Marblehead", "#00BFC4", "#F8766D")) %>%
  ggplot(aes(y = Municipality, x = per_capita_receipts, fill = FillCol)) +
  geom_col() +
  theme_minimal() +
   labs(title = "Per Capita Tax Receipts",
        y = "Municipality",
        x = "Taxes Per Capita") +
  geom_text(aes(label = scales::label_dollar(accuracy=1)(per_capita_receipts)), hjust = -0.1, size = 3)+
  guides(fill="none")+
  coord_cartesian(xlim = c(0, 8000)) +
  theme(plot.caption = element_text(color = "white"),axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(), 
        axis.title=element_blank(),
        axis.text=element_text(size=10),
        title=element_text(size=10))
plot8
```
```{r}
comp2 <- plot_grid(plot8, plot7, align = "h")
ggsave(here("figures", "taxsalpeers.png"), comp2, width = 6, height = 4, units = "in")
comp2
```

